Conditional selectivity for Hexastores

### 2011-07-11 conditional/trunk/selectivity-hexastore.tex
TODOs and status
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@7 43ef1190-90fd-4dd7-a769-d0121f02b80e
Add more todos and sel-equations
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@8 43ef1190-90fd-4dd7-a769-d0121f02b80e

\newcommand{\rdfnode}[1]{\texttt{#1}}

@@

@@

Several authors have studied the use of selectivity estimation for
optimizing SPARQL query execution plans, both in the context of Basic
Graph Patterns \cite{Stocker:2008:SBG:1367497.1367578} and in more
general cases. % TODO more references. OptARQ?

@@

Previous authors have taken selectivity as analogous to the
probability of matching a SPARQL node or triple with a node or triple
in the data. For a triple, $<S, P, O>$ the selectivity is the
intersection of the selectivity of each node.

@@

Analogously, the conditional selectivity for a triple $T$ can be written as 
\begin{equation}
sel(T) = sel(S \cap P \cap O) = sel(O | S \cap P) \cdot sel(P | S)
\cdot sel(S) . 
\end{equation}
We note that intersection is commutative, thus similar
relations can be shown for each permutation of nodes.

@@

% TODO example?

@@

Now, given a query, we shall explain how the selectivity can be
computed exactly rather than being estimated. First, we define a
function $cm(x)$ which returns the number of matches of a given SPARQL
node $x$. % TODO: needs to be a lot more precise

>>>>

\newcommand{\rdfnode}[1]{\texttt{#1}}
\newcommand{\todo}[1]{\tiny{TODO}\normalsize\footnote{\textbf{TODO:}~#1}}

@@

\section*{Status of this document}

This is currently at braindump status. Do not distribute without the
permission of the author.

@@

Several authors have studied the use of selectivity estimation for
optimizing SPARQL query execution plans, both in the context of Basic
Graph Patterns \cite{Stocker:2008:SBG:1367497.1367578} and in more
general cases. \todo{more references. OptARQ?}

@@

Previous authors have taken selectivity as analogous to the
probability of matching a SPARQL node\todo{terminology, (SPARQL nodes
  vs. RDF nodes) could be better} or triple with a node or triple
in the data. For a triple, $<S, P, O>$ the selectivity is the
intersection of the selectivity of each node.

@@

Analogously\todo{anybody good at arguing this mathematically?}, the conditional selectivity for a triple $T$ can be written as 
\begin{equation}
sel(T) = sel(S \cap P \cap O) = sel(O | S \cap P) \cdot sel(P | S)
\cdot sel(S) . 
\end{equation}
We note that intersection is commutative, thus similar
relations can be shown for each permutation of nodes.

@@

\todo{example?}

@@

Now, given a query, we shall explain how the selectivity can be
computed exactly rather than being estimated. First, we define a
function $cm(x)$ which returns the number of matches of a given SPARQL
node $x$. \todo{needs to be a lot more precise}

Then, 
\begin{equation}
sel(S) = \frac{cm(s_i)}{i} ,
\end{equation}
\begin{equation}
sel(P) = \frac{cm(p_{n_i}^i)}{n_i}
\end{equation}
and
\begin{equation}
sel(O) = \frac{cm(o_{k_{i,n_i}}^{i,n_i})}{k_{i,n_i}}
\end{equation}
